---
title: "Integration of scRNA-seq and Lipidomics Data - Mecp2 Project"
author: "Kari Neier"
date: "9/14/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Background

This experiment is designed to characterize transcriptomic profiles of Mecp2-e1 mutant vs. wild-type mice in the brain at the single-cell level. In addition, we also have lipidomic data from brain tissue of these same mice, and are interested in determining the following:

1) Which cell types have the strongest associations between gene expression and lipids that were differentially abundant in Mecp2-e1 mutants vs. wild-types?
2) Which genes are associated with brain lipids in diferent cell types? Are these genes also differentially expressed in Mecp2-e1 mutant vs. wild-types?

### Approach

Use limma to evaluate associations between lipids and gene expression, stratified by cell type. 

### The Data

- Labeled Lipids from an untargeted lipidomic analysis of brain tissue from male Mecp2-e1 mutants and wild-type littermate controls
- scRNA-seq data generated by Osman with data processed and cells labeled 

## Setting up

Loading packages, setting working directory, loading in the Seurat Object that was generated by Osman, and loading brain lipidomics data

```{r setting_up}
library(Seurat)
library(limma)
library(Glimma)

setwd("/Users/karineier/Documents/Mecp2/scRNA-seq")

load("/Users/karineier/Documents/Mecp2/scRNA-seq/Male_Cortex_labeled/rett_P30_with_labels_proportions.rda")

brain_lipids = read.csv("/Users/karineier/Documents/Mecp2/Brain_Lipidomics/Males_Brain_semiquant_data_normalized.csv")

brain_lipids_stats = read.csv("/Users/karineier/Documents/Mecp2/Brain_Lipidomics/Males_Brain_Lipidomics_Semiquant_t_test_annotated") # contains p-values and FDRs for brain lipids in males

```

## Kari you are here

You can also embed plots, for example:

```{r pressure, echo=FALSE}
plot(pressure)
```

Note that the `echo = FALSE` parameter was added to the code chunk to prevent printing of the R code that generated the plot.
